---
title: "Wendy"
output: html_document
---

```{r}
library('data.table')

# Set working directory to the directory containing this file and the AirlineDelays directory.
#setwd("/Users/wanxinbai/Documents/480-data science foundation/project/airlines")
setwd("/home/student/container-data/project")

## you can use fread first and then save the data as a RDdata, so you can just load data everytime you need it.

#air = fread('airlines.csv')
#save(air, file = 'air.RDdata')
load('air.RDdata')

#read in airports dataset and merge with origin airports
#airports <- read.csv("airports.csv")

#read for merging with destination airports
#dairports <- read.csv("airports.csv")
#names(dairports) <- c("diata","dairport", "dcity","dstate", "dcountry", "dlat", "dlong")

#merge origin airports
#PortLine <- merge(air, airports, by.x = 'Origin', by.y = 'iata')

#merge destination airports
#TPortLine <- merge (PortLine, dairports,  by.x = 'Dest', by.y = 'diata')

#merge with plane data and get alldata with 11210931 obs and 49 variables
#library(plyr)
#plane <- read.csv("plane-data.csv")
#plane <- rename(plane, c("year" = "plane_year"))
#alldata <- merge (TPortLine, plane ,by.x = 'TailNum', by.y = 'tailnum', all.x = T)
#save(alldata, file = 'alldata.RDdata')
load('alldata.RDdata')
alldata <- as.data.frame(alldata)
```


```{r}
##average delay rate for each hour in two years
#hourdelay <- read.csv("day.csv")
#head(air)
#names(hourdelay) <- c("hour","ave_delay")
#save(hourdelay, file = 'hourdelay.RDdata')

```


###pacakge preparing 

```{r,echo=FALSE,message=FALSE, warning=FALSE, results='hide'}
installIfNeeded = function(cliblist){
  libsNeeded = cliblist
  libsNeeded = libsNeeded[!(libsNeeded %in% installed.packages()[,"Package"])]
  if(length(libsNeeded)>0) install.packages(libsNeeded)
}

installIfNeeded("treemap")

# load the treemap package
library('treemap')

# Any package that is required by the script below is given here:
inst_pkgs = load_pkgs =  c("ggplot2","ggplot2movies", "dplyr","babynames","data.table","Rcpp")
inst_pkgs = inst_pkgs[!(inst_pkgs %in% installed.packages()[,"Package"])]
if(length(inst_pkgs)) install.packages(inst_pkgs)

git_pkgs = git_pkgs_load = c("streamgraph","DT")

git_pkgs = git_pkgs[!(git_pkgs %in% installed.packages()[,"Package"])]

if(length(git_pkgs)){
  library(devtools)
  install_github('rstudio/DT')
  install_github('hrbrmstr/streamgraph')
}

load_pkgs = c(load_pkgs, git_pkgs_load)

# Dynamically load packages
pkgs_loaded = lapply(load_pkgs, require, character.only=T)

```

### package for circle_diagrams

```{r}
inst_pkgs = load_pkgs =  c("circlize")
inst_pkgs = inst_pkgs[!(inst_pkgs %in% installed.packages()[,"Package"])]
if(length(inst_pkgs)) install.packages(inst_pkgs)

# Dynamically load packages
pkgs_loaded = lapply(load_pkgs, require, character.only=T)

# Chord graphs are fantastic at showing the links or relationships that exist between categories.
circos.clear()
```




#### generate new variables, if the flight delayed or not

```{r}
alldata$Delay = ifelse(alldata$ArrDelay > 0, 1, 0)
```


### Variable Exploration

#### unique carrier

#### for two years

```{r}
treemap(air,
        index=c("UniqueCarrier"),
        vSize="Delay",
        vColor="Delay",
        type="value")#actual value or percentage value
#if we are using propotion

count <- table(air$UniqueCarrier) #flight count for each carrier
delaycount <- table(air[which(air$Delay == 1)]$UniqueCarrier) # delay count for each carrier
delayprop <- delaycount/count
delayprop <- as.data.frame(delayprop)
names(delayprop)[1] <- 'UniqueCarrier'
sortedelayprop <- delayprop[order(delayprop$Freq),]
sortedelayprop

summary(sortedelayprop$Freq)


#total time of delay
treemap(delayprop,
        index=c("UniqueCarrier"),
        vSize="Freq",
        vColor="Freq",
        type="value")#actual value or percentage value


treemap(air,
        index=c("UniqueCarrier"),
        vSize="ArrDelay",
        vColor="ArrDelay",
        type="value")#actual value or percentage value

#average delay time for each carrier
DTsum <- aggregate(ArrDelay ~ UniqueCarrier, air[which(air$ArrDelay > 0)], sum)
DTsum[2] <- DTsum[2]/count
sortedDTsum <- DTsum[order(DTsum$ArrDelay),]
sortedDTsum
summary(sortedDTsum$ArrDelay)

treemap(DTsum,
        index=c("UniqueCarrier"),
        vSize="ArrDelay",
        vColor="ArrDelay",
        type="value")#actual value or percentage value

```


####For year 1999

```{r}
air99 = air[Year == 1999, ]
treemap(air99,
        index=c("UniqueCarrier"),
        vSize="Delay",
        vColor="Delay",
        type="value")#actual value or percentage value
#if we are using propotion

count <- table(air99$UniqueCarrier) #flight count for each carrier
delaycount <- table(air99[which(air99$Delay == 1)]$UniqueCarrier) # delay count for each carrier
delayprop <- delaycount/count
delayprop <- as.data.frame(delayprop)
names(delayprop)[1] <- 'UniqueCarrier'
sortedelayprop <- delayprop[order(delayprop$Freq),]
sortedelayprop

summary(sortedelayprop$Freq)


#total time of delay
treemap(delayprop,
        index=c("UniqueCarrier"),
        vSize="Freq",
        vColor="Freq",
        type="value")#actual value or percentage value


treemap(air99,
        index=c("UniqueCarrier"),
        vSize="ArrDelay",
        vColor="ArrDelay",
        type="value")#actual value or percentage value

#average delay time for each carrier
DTsum <- aggregate(ArrDelay ~ UniqueCarrier, air99[which(air99$ArrDelay > 0)], sum)
DTsum[2] <- DTsum[2]/count
sortedDTsum <- DTsum[order(DTsum$ArrDelay),]
sortedDTsum
summary(sortedDTsum$ArrDelay)

treemap(DTsum,
        index=c("UniqueCarrier"),
        vSize="ArrDelay",
        vColor="ArrDelay",
        type="value")#actual value or percentage value

```


####For year 2000

```{r}
air00 = air[Year == 2000, ]
treemap(air00,
        index=c("UniqueCarrier"),
        vSize="Delay",
        vColor="Delay",
        type="value")#actual value or percentage value
#if we are using propotion

count <- table(air00$UniqueCarrier) #flight count for each carrier
delaycount <- table(air00[which(air00$Delay == 1)]$UniqueCarrier) # delay count for each carrier
delayprop <- delaycount/count
delayprop <- as.data.frame(delayprop)
names(delayprop)[1] <- 'UniqueCarrier'
sortedelayprop <- delayprop[order(delayprop$Freq),]
sortedelayprop

summary(sortedelayprop$Freq)


#total time of delay
treemap(delayprop,
        index=c("UniqueCarrier"),
        vSize="Freq",
        vColor="Freq",
        type="value")#actual value or percentage value


treemap(air00,
        index=c("UniqueCarrier"),
        vSize="ArrDelay",
        vColor="ArrDelay",
        type="value")#actual value or percentage value

#average delay time for each carrier
DTsum <- aggregate(ArrDelay ~ UniqueCarrier, air00[which(air00$ArrDelay > 0)], sum)
DTsum[2] <- DTsum[2]/count
sortedDTsum <- DTsum[order(DTsum$ArrDelay),]
sortedDTsum
summary(sortedDTsum$ArrDelay)

treemap(DTsum,
        index=c("UniqueCarrier"),
        vSize="ArrDelay",
        vColor="ArrDelay",
        type="value")#actual value or percentage value

```


##DepDelay & Origin airport

### For two years

```{r}
treemap(air,
        index=c("Origin"),
        vSize="Delay",
        vColor="Delay",
        type="value")#actual value or percentage value
#if we are using propotion
count <- table(air$Origin) #flight count for each origin airport
delaycount <- table(air[which(air$Delay == 1)]$Origin) # delay count for each origin airport
delayprop <- delaycount/count
delayprop <- as.data.frame(delayprop)
names(delayprop)[1] <- 'Origin'
sortedelayprop <- delayprop[order(delayprop$Freq),]
sortedelayprop
summary(sortedelayprop$Freq)



#total time of delay
treemap(delayprop,
        index=c("Origin"),
        vSize="Freq",
        vColor="Freq",
        type="value")#actual value or percentage value

#average delay time for each origin airport
DTsum <- aggregate(ArrDelay ~ Origin, air[which(air$ArrDelay > 0)], sum)
DTsum[2] <- DTsum[2]/count
sortedDTsum <- DTsum[order(DTsum$ArrDelay),]
sortedDTsum
summary(sortedDTsum$ArrDelay)

treemap(DTsum,
        index=c("Origin"),
        vSize="ArrDelay",
        vColor="ArrDelay",
        type="value")#actual value or percentage value

```

### For 1999

```{r}
treemap(air99,
        index=c("Origin"),
        vSize="Delay",
        vColor="Delay",
        type="value")#actual value or percentage value
#if we are using propotion
count <- table(air99$Origin) #flight count for each origin airport
delaycount <- table(air99[which(air99$Delay == 1)]$Origin) # delay count for each origin airport
delayprop <- delaycount/count
delayprop <- as.data.frame(delayprop)
names(delayprop)[1] <- 'Origin'
sortedelayprop <- delayprop[order(delayprop$Freq),]
sortedelayprop
summary(sortedelayprop$Freq)



#total time of delay
treemap(delayprop,
        index=c("Origin"),
        vSize="Freq",
        vColor="Freq",
        type="value")#actual value or percentage value

#average delay time for each origin airport
DTsum <- aggregate(ArrDelay ~ Origin, air99[which(air$ArrDelay > 0)], sum)
DTsum[2] <- DTsum[2]/count
sortedDTsum <- DTsum[order(DTsum$ArrDelay),]
sortedDTsum
summary(sortedDTsum$ArrDelay)

treemap(DTsum,
        index=c("Origin"),
        vSize="ArrDelay",
        vColor="ArrDelay",
        type="value")#actual value or percentage value

```

### For 2000

```{r}
treemap(air00,
        index=c("Dest"),
        vSize="Delay",
        vColor="Delay",
        type="value")#actual value or percentage value
#if we are using propotion

count <- table(air00$Dest) #flight count for each origin airport
delaycount <- table(air00[which(air00$Delay == 1)]$Dest) # delay count for each origin airport
delayprop <- delaycount/count
delayprop <- as.data.frame(delayprop)
names(delayprop)[1] <- 'Dest'
sortedelayprop <- delayprop[order(delayprop$Freq),]
sortedelayprop
summary(sortedelayprop$Freq)


#total time of delay
treemap(delayprop,
        index=c("Dest"),
        vSize="Freq",
        vColor="Freq",
        type="value")#actual value or percentage value

#average delay time for each origin airport
DTsum <- aggregate(ArrDelay ~ Dest, air00[which(air00$ArrDelay > 0)], sum)
DTsum[2] <- DTsum[2]/count
sortedDTsum <- DTsum[order(DTsum$ArrDelay),]
sortedDTsum
summary(sortedDTsum$ArrDelay)

treemap(DTsum,
        index=c("Dest"),
        vSize="ArrDelay",
        vColor="ArrDelay",
        type="value")#actual value or percentage value

```



###plane information



###Timeline
####Year and arrdelay

```{r, eval=FALSE}
yeardata <- air[,c("Year","Month","ArrDelay")]
yeardata <- as.data.frame(yeardata)
yeardata$Year <- as.factor(yeardata$Year)
yeardata$Month <- as.factor(yeardata$Month)

# Stacked Density Plot
library(ggplot2)
ggplot(yeardata, aes(x=ArrDelay)) + xlim(-50,100) +
  geom_density(aes(group=Year, colour=Year, fill=Year, alpha = 0.4 ), position="stack")

ggplot(yeardata, aes(x=ArrDelay)) + xlim(-50,100) +
  geom_density(aes(group=Year, colour=Year, fill=Year, alpha = 0.4 ))

```
 
Almost symetric between 0, but thicker  tails  when arrdelay > 0. And almost the same for two years.

```{r, eval=FALSE}
ggplot(yeardata, aes(x=ArrDelay)) + xlim(-50,50) +
  geom_density(aes(group=Month, colour=Month, fill=Month, alpha = 0.4 ), position = "stack")

```


#DepDelay for each month

```{r, eval =FALSE}

library(foreach)
library(streamgraph)

yeardata <- as.data.frame(yeardata)
data <- aggregate(ArrDelay ~ Year + Month, yeardata[yeardata$ArrDelay > 0,], sum)
data$Month <- paste("2012",data$Month,"01",sep = '-')
# years <- split(1:nrow(x), air[,"Year"])
months <- split(1:nrow(air), air[,"Month"])
# DOMs <- split(1:nrow(x), x[,"DayofMonth"]) 
# DOWs <- split(1:nrow(x), x[,"DayOfWeek"])
# origin <- split(1:nrow(x), x[,"Origin"])
gyear <- group_by(yeardata, Year, Month)
gyear <- ungroup(gyear)
streamgraph(data, "Year", "ArrDelay", "Month", interactive=TRUE) %>%
  sg_fill_brewer("PuOr")

#sg_axis_x(1, "Month", "%M") %>%

streamgraph(data, "Year", "ArrDelay", "Month") %>%
  sg_fill_brewer("Spectral") %>%
  sg_axis_x(tick_units = "Month", tick_interval = 1)



numflights <- foreach(month = months, .combine = cbind) %do% {
  flights <- air[month,]
  count99 <- length(which(flights[,"Year"] == "1999"))
  count00 <- length(which(flights[,"Year"] == "2000"))
  rbind(count99, count00)
}
dq <- melt(numflights)
names(dq) <- c("year", "month", "count")
# conver month column to date format for streamgraph
dq$month <- paste("2012", dq$month, "01", sep = "-")


streamgraph(dq, "", "count","month", interactive=TRUE)



```



```{r, eval=FALSE}
treemap(air,
        index=c("Month"),
        vSize="Delay",
        vColor="Delay",
        type="value")#actual value or percentage value
#if we are using propotion
streamgraph(air, "Year","ArrDelay", "Month", interactive=TRUE) %>%
  sg_axis_x("Month", "%Y") %>%
  sg_fill_brewer("PuOr")


count <- table(air$Month) #flight count for each month
delaycount <- table(air[which(air$Delay == 1)]$Month) # delay count for each month
delayprop <- delaycount/count
delayprop <- as.data.frame(delayprop)
names(delayprop)[1] <- 'Month'
sortedelayprop <- delayprop[order(delayprop$Freq),]
summary(sortedelayprop$Freq)


#total time of delay
treemap(delayprop,
        index=c("Month"),
        vSize="Freq",
        vColor="Freq",
        type="value")#actual value or percentage value

#average delay time for each origin airport
DTsum <- aggregate(ArrDelay ~ Dest, air[which(air$ArrDelay > 0)], sum)
DTsum[2] <- DTsum[2]/count
sortedDTsum <- DTsum[order(DTsum$ArrDelay),]
summary(sortedDTsum$ArrDelay)

treemap(DTsum,
        index=c("Dest"),
        vSize="ArrDelay",
        vColor="ArrDelay",
        type="value")#actual value or percentage value

```



###numerica variables

```{r,eval = FALSE}
pairs(air[,c("CRSDepTime","CRSArrTime","ActualElapsedTime","CRSElapsedTime","AirTime","ArrDelay","DepDelay","Distance", "TaxiIn","TaxiOut")],pch = ".")
```


```{r, eval=FALSE}
pairs(air[,c("CRSDepTime","CRSArrTime","ArrDelay","DepDelay")],pch = ".")
pairs(air[,c("ActualElapsedTime","CRSElapsedTime","ArrDelay","DepDelay")],pch = ".")
pairs(air[,c("AirTime","ArrDelay","DepDelay","Distance")],pch = ".")
pairs(air[,c("ArrDelay","DepDelay", "TaxiIn","TaxiOut")],pch = ".")
```



### AirTime and Distance with Delay

```{r}
library(ggplot2)
noair <- na.omit(air[,c("Delay","Distance","AirTime","ArrDelay")])
x_axis = "Delay"
y_axis = "Distance"
color  = "AirTime"
ggplot(noair, aes_string(x=x_axis, y=y_axis, color=color)) +
  geom_point(position=position_jitter(w=0.0, h=0.4)) +
  theme_light(base_size=20) +
  xlab("Delay Or Not") +
  ylab("Distance ") +
  scale_colour_gradientn("AirTime", colours=c("#5e4fa2", "#3288bd", "#66c2a5", "#abdda4", "#e6f598", "#fee08b", "#fdae61", "#f46d43", "#d53e4f", "#9e0142")) +
  ggtitle("Airtime and Distance") +
  theme(plot.title=element_text(size=18))
```


### AirTime and Distance with delaytime

```{r}
x_axis = "AirTime"
y_axis = "Distance"
color  = "ArrDelay"
ggplot(noair, aes_string(x=x_axis, y=y_axis, color=color)) +
  geom_point(position=position_jitter(w=0.0, h=0.4)) +
  theme_light(base_size=20) +
  xlab("AirTime") +
  ylab("Distance ") +
  scale_colour_gradientn("ArrDelay", colours=c("#5e4fa2", "#3288bd", "#66c2a5", "#abdda4", "#e6f598", "#fee08b", "#fdae61", "#f46d43", "#d53e4f", "#9e0142")) +
  ggtitle("Airtime and Distance") +
  theme(plot.title=element_text(size=18))
```


#top_n(10, n)



