---
title: "Manu2000"
output: html_document
author: Wanxin Bai
---

```{r}
setwd("/home/student/container-data/project")
#load('air.RDdata')
#air <- as.data.frame(air)
load('alldata.RDdata')
alldata <- as.data.frame(alldata)
```


#### generate new variables, if the flight delayed or not

```{r}
alldata$Delay = ifelse(alldata$ArrDelay > 0, 1, 0)
```


###pacakge preparing 

```{r,echo=FALSE,message=FALSE, warning=FALSE, results='hide'}
library('bit64')
installIfNeeded = function(cliblist){
  libsNeeded = cliblist
  libsNeeded = libsNeeded[!(libsNeeded %in% installed.packages()[,"Package"])]
  if(length(libsNeeded)>0) install.packages(libsNeeded)
}

installIfNeeded("treemap")

# load the treemap package
library('treemap')

# Any package that is required by the script below is given here:
inst_pkgs = load_pkgs =  c("ggplot2","ggplot2movies", "dplyr","babynames","data.table","Rcpp")
inst_pkgs = inst_pkgs[!(inst_pkgs %in% installed.packages()[,"Package"])]
if(length(inst_pkgs)) install.packages(inst_pkgs)

git_pkgs = git_pkgs_load = c("streamgraph","DT")

git_pkgs = git_pkgs[!(git_pkgs %in% installed.packages()[,"Package"])]

if(length(git_pkgs)){
  library(devtools)
  install_github('rstudio/DT')
  install_github('hrbrmstr/streamgraph')
}

load_pkgs = c(load_pkgs, git_pkgs_load)

# Dynamically load packages
pkgs_loaded = lapply(load_pkgs, require, character.only=T)

```




####For year 2000

```{r}
alldata00 = alldata[alldata$Year == 2000, ]
treemap(alldata00,
        index=c("manufacturer"),
        vSize="Delay",
        vColor="Delay",
        type="value")
```

MCDONNELL DOUGLAS had a lower delay count in 2000.


####if we are using propotion

```{r}

count <- table(alldata00$manufacturer) #flight count for each manufacturer
delaycount <- table(alldata00[which(alldata00$Delay == 1),"manufacturer"]) # delay count for each carrier
delayprop <- delaycount/count
delayprop <- as.data.frame(delayprop)
names(delayprop)[1] <- 'manufacturer'
sortedelayprop <- delayprop[order(delayprop$Freq, decreasing = T),]
sortedelayprop[sortedelayprop$Freq != 'NaN',]

```

BAUMAN RANDY had a really bad performance this year. So did AIRBUS.

```{r}
summary(sortedelayprop$Freq)
```


#### treemap for delay rate for each manufacturer

```{r}
treemap(delayprop,
        index=c("manufacturer"),
        vSize="Freq",
        vColor="Freq",
        type="value")


treemap(alldata00,
        index=c("manufacturer"),
        vSize="ArrDelay",
        vColor="ArrDelay",
        type="value")
```

AIRBUS 's delay time increased for this year.


```{r}
#average delay time for each carrier
DTsum <- aggregate(ArrDelay ~ manufacturer, alldata00[which(alldata00$ArrDelay > 0),], sum)
DTsum[2] <- DTsum[2]/count
DTsum <- DTsum[DTsum$ArrDelay < 100000,]
sortedDTsum <- DTsum[order(DTsum$ArrDelay),]
sortedDTsum
summary(sortedDTsum$ArrDelay)
```

No big changes for the rank of average delay rate for each manufacturers and still had some extremely high delay time for several companies.

```{r}
treemap(DTsum,
        index=c("manufacturer"),
        vSize="ArrDelay",
        vColor="ArrDelay",
        type="value")

```

